Lattice Boltzmann simulations of magnetically responsive\\ ellipsoidal particle stabilized emulsions

### 2025-03-19 chapters/paper3.tex
Changes 1. Completed rewrite of aim 3 chapter

I show in Figure \ref{fig:particles_interface_prop_shear} that when the magnetic field is kept the same between the template and during shear, the
proportion of particles on the interface ($n_{int}$) stays approximately the same. However when changing the applied field strength, $n_{int}$ more 
sensitive to shear. Here I also characterize that increasing the applied field strength reduces $n_{int}$ more than decreasing it. To characterize 
the role of the interface properties and the impact of particle ejection and tilting of particles out of the interface,the next section aims to 
uncover the microstructural implications of these effects.

% This is consistent
% with the amount of coarsening observed for each case seen in Figure \ref{fig:domain_size_time_shear} where $\bar{B}:0 \rightarrow 1$ shows the largest
% change.

@@

Bonaccorso et al. identified that the spherical particle stabilizers in their bijel simulations oriented to the direction of shear before being
ejected from the interface. \cite{bonaccorso_shear_2020} This then led to the domains of the bijel to coalesce and eventually destruction of the
bicontinuous microstructure. \cite{bonaccorso_shear_2020} In our bijels, we also characterize this behavior upon application of shear.
In this section, I probe the effect of applied shear to the microstructure of magnetically
responsive bijels stabilized by ellipsoidal particles. To investigate the effect of the particle angle on the interfacial coverage of the bijel,
I plot the area averaged gaussian curvature over time. This parameter will allow insight into the evolution of the interface, namely the presence of
domain coarsening, coalscence or break-off events.

@@

Figure \ref{fig:gaussian_curvature_time_shear} reveals that the area averaged Gaussian curvature generally decreases in magnitude over time for all cases except 
$\bar{B}: 0 \to 1$. I also characterize that the curvature evolution of all systems are shear independent. $\bar{B}: 0 \to 1$ is
unique because particle rearrangements at the interface dominates the dynamics of the microstructure, not observed in all other
cases. From the time evolution of the curvature, I show that the same mechanism drives domain coarsening in all cases. Domain
coarsening is driven through a reduction in the interfacial area covered by the particle monolayer. From the curvature results characterized in 
Figure \ref{fig:gaussian_curvature_time_shear}, I saw that the coarsening is shear independent. This indicates that the surface area of the interface 
decreases which can occur due to changes in the interfacial orientation of particles at the interface or a reduction in the number of particles at the interface. 
I can characterize the former mechanism using the average interfacial angle between the particle and the interface $\langle \psi \rangle$.

>>>>

From the results in Figure \ref{fig:particles_interface_prop_shear}, the impact of the initial microstructure on the number of particles on the interface
differs for bijels stabilized by oblate and prolate ellipsoids. We split this analysis on processing history. While keeping the applied magnetic field
the same, we see little differences in the behavior of $n_{int}$, with their behavior corresponding to changes in $\langle \psi \rangle$. 
Looking at the case of $\bar{B}:0 \rightarrow 0$, there is little change in the number of particles at the interface upon application of shear. This implies that the 
particles remain strongly adsorbed on the
interface. The $\langle \psi \rangle$ results also show this as we see this parameter increasing slowly. This suggests that the particles remain firmly adsorbed
on the interface. Looking to the $\bar{B}:0 \rightarrow 1$ case, the number of particles at the interface drops while $\langle \psi \rangle$ increases. This
trend affects both particles which suggests that shear can pull particles out of the interface due to the decreased
interfacial adsorption energy the particle experiences when rotating out of the interface to orient to the applied field. The $\bar{B}:1 \rightarrow 1$ and 
$\bar{B}:1 \rightarrow 0$ have strong dependences on particle morphology. While an increase in $\langle \psi \rangle$ can be seen in both processing histories 
for both particle morphologies, a decrease in $n_{int}$ can be seen for prolate particles with processing history $\bar{B}:1 \rightarrow 0$ and for 
oblate particles with processing history $\bar{B}:1 \rightarrow 1$. We visualize $\langle \psi \rangle$ at the initial and final timestep for bijels stabilized with
oblate and prolate particles with processing $\bar{B}:1 \rightarrow 0$.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.4]{../figures/results/paper3/psi_concat_startB-1_endB-0.png} 
    \caption{Visualization of the particle monolayer of bijels stabilized by prolate and oblate ellipsoidal particles with processing history $\bar{B}: 1 \rightarrow 0$
             undergoing an applied shear of $Ca_s = 10^{-6}$. Particles with a deeper red shade have a larger angle to the interface. The top and bottom row show the 
             configuration of the particles at the initial and final timestep respectively.} 
    \label{fig:particle_viz_psi} 
\end{figure}

Figure \ref{fig:particle_viz_psi} shows that the increase in $\langle \psi \rangle$ seen in Figure \ref{fig:interface_angle_shear} occurs due to different
reasons for both particle morphologies. For oblate particles, this is caused by particles orienting such that their symmetry axis is parallel to the interface
while prolate particles experienced interparticle forces induced by the applied shear, resulting in particles being ejected from the interface when the magnetic 
field is switched off. When keeping the magnetic field on, the oblate particles are forced to stay in sub-optimal particle orientations to the interface, resuling
in ejection of particles from the interface. The presence of the magnetic field reduces shear induced interparticle interactions owing to the ordering of
particles to the applied field, resulting in the mostly unchanged $n_{int}$ values. A increase in $\langle \psi \rangle$ and decrease in $n_{int}$ indicates domain
can indicate coarsening of the fluid domains as the interfacial area is reduced. In the next section, I investigate the the role of the interface properties and the 
impact of particle ejection and tilting of particles out of the interface on the microstructure of bijels.

@@

Bonaccorso et al. identified that the spherical particle stabilizers in their bijel simulations oriented to the direction of shear before being
ejected from the interface. \cite{bonaccorso_shear_2020} This then led to the domains of the bijel to coalesce and eventually destruction of the
bicontinuous microstructure. \cite{bonaccorso_shear_2020} In our bijels, we characterize particle morphology and initial microstructure specific 
properties of the particle monolayer upon application of shear. 
In this section, I probe the effect of applied shear to the microstructure of magnetically
responsive bijels stabilized by ellipsoidal particles. To investigate the effect of the particle angle and the interfacial coverage of the bijel on the
microstructural evolution of the interface, I plot the area averaged gaussian curvature over time. This parameter will allow insight into the evolution 
of the interface, namely the presence of domain coarsening, coalscence or break-off events.

@@

Figure \ref{fig:gaussian_curvature_time_shear} reveals that the area averaged Gaussian curvature generally decreases in magnitude over time for all cases except 
$\bar{B}: 0 \to 1$ indicating that the increase in $\langle \psi \rangle$ causes domain coarsening of the bijel. $\bar{B}: 0 \to 1$ is
unique because particle rearrangements at the interface dominates the dynamics of the microstructure, not observed in all other
cases. From the time evolution of the curvature, the same mechanism drives coarsening in all cases. From the results in Figures \ref{fig:particles_interface_prop_shear} 
and \ref{fig:interface_angle_shear}, domain coarsening is driven through a reduction in the interfacial area covered by the particle monolayer. 
From the curvature results characterized in Figure \ref{fig:gaussian_curvature_time_shear}, the area averaged curvature coarsening is shear independent. This is further
evidence that the mechanism of coarsening is identical in each case.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/anisotropy_compare.png} 
    \caption{Comparisons of the time evolution of the directional domain sizes that increase upon application of a magnetic field for bijels stabilized 
             by ellipsoidal particles undergoing various shear rates under four processes. The directional domain size increases in all cases except for when the
             field is switched off.} 
    \label{fig:domain_size_aniso_time_shear} 
\end{figure}

From Figure \ref{fig:domain_size_aniso_time_shear}, the directional domain sizes in the orientations with the maximum change were characterized. For oblate and prolate 
particles, this was found to be $L_{\perp}$ and $L_{\parallel}$ respectively. In all cases except for $\bar{B}: 1 \to 0$, this length scale increases matching the expected
trend that increasing $\langle \psi \rangle$ matches domain coarsening. The decrease in these length scales for $\bar{B}: 1 \to 0$ suggest that the shear induced
reordering of interfacial particles after application of a field affects the evolution of the microstructure. For both particles, this is due to particle mediated
shear forces causing domains to coarsen more in specific directions. From the snapshots of the particle arrangement in Figure \ref{fig:particle_viz_psi} the particles were
shown to reorient at the interface. This process did not occur for $\bar{B}: 1 \to 1$ as the presence of magnetic fields acted to keep particles aligned a specific
direction.

In this chapter, I investigated the rheology of bijels stabilized by ellipsoidal particles under shear, focusing on the influence of initial microstructure and the 
application of a constant magnetic field on their shear response. The shear behavior was characterized using a Herschel-Bulkley fitting function to extract the flow index, 
consistency index, and yield stress. My analysis confirmed that the shear-thinning nature of bijels remains robust, regardless of particle morphology or applied magnetic 
field. However, the flow index and yield stress trends were strongly dependent on these factors.  
I found that increased particle ordering led to distinct yield stress behaviors: oblate particles resulted in a lower yield stress, while prolate particles exhibited a 
higher yield stress. When examining the properties of the particle monolayer, I identified a clear impact of particle morphology on interparticle interactions at the 
interface. In the absence of a magnetic field, both particle monolayers remained adsorbed at the interface. However, when a magnetic field was initially applied, the 
behavior diverged significantly oblate particles flipped out of the interface to minimize interfacial energy, whereas prolate particles tilted out. These differences 
directly influenced the shear response, altering how the surrounding fluid flows around the particles and leading to the observed non-Newtonian behavior and yield 
stress characteristics.  
Furthermore, I investigated the microstructural effects of shear and magnetic field-induced particle ordering and found that coarsening was the primary driver of 
domain size evolution in sheared bijels. This coarsening was facilitated by a reduction in interfacial particle coverage. When analyzing the anisotropic domain size evolution, 
I observed that in most cases, the direction of greatest structural change aligned with expected trends—except when the magnetic field was switched off. This deviation is 
attributed to differences in particle monolayer evolution, depending on whether the bijel was formed with or without an initially applied field.
